---
title: "Data Dictionary"
author: "Jennifer Collister"
date: "19/02/2020"
output: word_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(knitr)
```

```{r loaddata, include=FALSE}

data <- readRDS(file="K:\\TEU\\APOE on Dementia\\Data Management\\R_Dataframes_TLA\\38358\\Organised\\Hypertension\\Neo\\HTN_excl.rds")

```

# Core analysis

## Core endpoints

* hypertension awareness
* hypertension treatment
* hypertension control

## Explanatory variables of interest

* Ethnicity

## Covariates

* Age
* Gender
* Education
* BMI



```{r coreanalysis, include=FALSE}
# (varname, varlabel, description, vartype, UKBfield, response, n)
core <- list("aware"=list(), "treated"=list(), "controlled"=list(), 
             "ethnicity"=list(), 
             "age"=list(), "gender"=list(), "education"=list(), "BMI"=list())

# Hypertension awareness
core[["aware"]] <- list("varname"="aware", 
              "varlabel"="Participant awareness of hypertension", 
              "description"="If the participant self-reported a diagnosis of hypertension in the touchscreen questionnaire or verbal interview, or reported that they were taking BP medication", 
              "vartype"="Boolean", 
              "UKBfield"="6150, 6153, 6177, 20002, 20003", 
              "response"=c(paste0("FALSE: ", levels(data$aware_)[1]), paste0("TRUE: ", levels(data$aware_)[2])),
              "n"=table(data$aware)
              )

# Hypertension treatment
core[["treated"]] <- list("varname"="treated", 
              "varlabel"="Participant treatment of hypertension", 
              "description"="Among participants who were aware of their hypertension, if they self-reported taking BP meds or the list of medication they provided in verbal interview was found to include a standard hypertensive treatment regimen then they were considered to be treated for hypertension", 
              "vartype"="Boolean", 
              "UKBfield"="6153, 6177, 20003", 
              "response"=c(paste0("FALSE: ", levels(data$treated_)[1]), paste0("TRUE: ", levels(data$treated_)[2])),
              "n"=table(data$treated)
              )

# Hypertension control
core[["controlled"]] <- list("varname"="controlled", 
              "varlabel"="Participant control of hypertension", 
              "description"="Among participants who were taking treatment for hypertension, if their mean measured BP at baseline assessment was < 140/90 then they was considered to be successfully controlling their BP", 
              "vartype"="Boolean", 
              "UKBfield"="93, 94, 4079, 4080", 
              "response"=c(paste0("FALSE: ", levels(data$controlled_)[1]), paste0("TRUE: ", levels(data$controlled_)[2])),
              "n"=table(data$controlled)
              )

# Ethnicity
core[["ethnicity"]] <- list("varname"="eth_group", 
              "varlabel"="Ethnic group", 
              "description"="Broad ethnic category of individuals", 
              "vartype"="Character", 
              "UKBfield"="21000", 
              "response"=c(levels(data$eth_group)),
              "n"=table(data$eth_group)
              )

# Age
core[["age"]] <- list("varname"="age", 
              "varlabel"="Age at baseline assessment (years)", 
              "description"="Age at baseline assessment, estimated from month and year of birth and date of attending recruitment centre", 
              "vartype"="Continuous numeric", 
              "UKBfield"="34, 52, 53", 
              "response"="Non-missing values",
              "n"=nrow(data[!is.na(data$age),])
              )

# Gender
core[["gender"]] <- list("varname"="gender", 
              "varlabel"="Gender", 
              "description"="Self-reported gender", 
              "vartype"="Character", 
              "UKBfield"="31", 
              "response"=c(levels(data$gender)),
              "n"=table(data$gender)
              )

# Education
core[["education"]] <- list("varname"="edu_highest", 
              "varlabel"="Highest achieved qualification", 
              "description"="Highest reported educational qualification, derived from list of qualifications reported in touchscreen questionnaire", 
              "vartype"="Character", 
              "UKBfield"="6138", 
              "response"=c(levels(data$edu_highest)),
              "n"=table(data$edu_highest)
              )

# BMI
core[["BMI"]] <- list("varname"="BSM_BMI", 
              "varlabel"="Body mass index at baseline assessment", 
              "description"="Derived by UKB from weight and height at baseline assessment", 
              "vartype"="Continuous numeric", 
              "UKBfield"="21001", 
              "response"="Non-missing values",
              "n"=nrow(data[!is.na(data$BSM_BMI),])
              )

coredf <- data.frame(do.call(rbind, core))
```


```{r tabulate_core, include=TRUE}

# kable(coredf)

kable(c(table(data$aware_), 
        table(data$treated_), 
        table(data$controlled_), 
        table(data$eth_group),
        "age"=nrow(data[!is.na(data$age),]),
        table(data$gender),
        table(data$edu_highest),
        "bmi"=nrow(data[!is.na(data$BSM_BMI),])
        )
      )
 
```



# Expanded analysis

# Explanatory variables

* Chronic co-morbidities
  * Number
  * Duration of diagnosis
  * Type
    * Asthma
    * COPD
    * Diabetes
    * Cancer
    * Depression

## Other covariates

* Sociodemographic characteristics
  * Country of birth
  * Household income
    * Continuous
    * Quintiles
  * Townsend score
  * Employment shift type
  * Occupation type
* Lifestyle behaviour
  * Tobacco
  * Harmful alcohol use
  * Physical activity
    * Continuous METS
    * > or <= 150 METs
  * Sleep duration
* Function
  * Reaction time
  * Handgrip strength
* Hypertension disease characteristics
  * Duration of diagnosis
  * Severity (Stage I, II, III)
  * Known history of stroke
    * Yes/No
    * Type
    * Duration of diagnosis
  * Known history of ischemic coronary disease
    * Yes/No
    * Type
    * Duration of diagnosis
  * Family history of hypertension
* Physical measurements
  * Waist circumference
  * BP1
  * BP2
* Baseline blood chemistry tests
  * Glucose
  * Total cholesterol
  * HDL cholesterol
  * LDL cholesterol
  * Triglycerides

```{r tabulate_extra, include=TRUE}

# kable(coredf)

kable(c(table(data$aware_), 
        table(data$treated_), 
        table(data$controlled_), 
        table(data$eth_group),
        "age"=nrow(data[!is.na(data$age),]),
        table(data$gender),
        "bmi"=nrow(data[!is.na(data$BSM_BMI),])
        )
      )
 
```





