---
title: "HTN PRS MACE prospective"
author:  "Jennifer Collister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  word_document:
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)
library(Publish)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)
library(png)
library(dplyr)
library(ggplot2)
library(survival)
library(pROC)

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(config$functions$JC))

#-------------------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r data-derivation, include=FALSE, cache=TRUE}
 
source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data){
  
  excl$initial <<- nrow(data)
  
  # Exclude those outside the 40-70 age range
  data <- data[data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  
  excl$agerange <<- nrow(data)

  # Exclude individuals with missing BP data
  # or missing answers to BP questions on touchscreen questionnaire
  data <- data[!is.na(data$TEU_BlP_SBP.avg),]
  data <- data[!is.na(data$TEU_BlP_DBP.avg),]
  data <- data[data$TEU_HMH_prevHTN != "Unanswered",]
  data <- data[data$TEU_HMH_Meds_BP != "Unanswered",]
  # Exclude participants who only had BP measured once
  data <- data[data$TEU_BlP_nSBP == 2 & data$TEU_BlP_nDBP == 2,]
  
  excl$BPmiss <<- nrow(data)

  # Exclude individuals with implausible BP data
  data <- data[data$TEU_BlP_SBP.avg >= 70 & data$TEU_BlP_SBP.avg <= 270,]
  data <- data[data$TEU_BlP_DBP.avg >= 50 & data$TEU_BlP_DBP.avg <= 150,]
  
  excl$BPimp <<- nrow(data)
  
  # Only keep those with genetic data
  data <- data[!is.na(data$TEU_BP_PRS),]
  
  excl$PRSmiss <<- nrow(data)

  # Exclude pregnant women ("Yes" or "Unsure")
  data <- data[is.na(data$VeI_PregnantNow) | data$VeI_PregnantNow == "No",]
  
  excl$pregnant <<- nrow(data)

  # Exclude individuals who have serious health conditions
  data <- data[is.na(data$TEU_VeI_seriouscomb),]

  # And individuals with cancer (except for skin cancer?)
  data <- data[is.na(data$TEU_VeI_cancer),]
  
  excl$seriouscomorb <<- nrow(data)
  
  # Any durations less than 0 are clearly errors (only 2)
  # Any diagnoses before the age of about 20 are probably due to other causes
  data$TEU_VeI_HTN_dur[data$TEU_VeI_HTN_dur < 0] <- NA
  # data[data$TEU_VeI_HTN_dur > data$TEU_BaC_AgeAtRec-20,]
  
  # Exclude those who were not hypertensive
  data <- data[data$TEU_evidenceHTN == TRUE & !is.na(data$TEU_evidenceHTN),]
  
  excl$evidenceHTN <<- nrow(data)
  
  # Exclude those who were not aware
  data <- data[data$TEU_awareHTN == TRUE & !is.na(data$TEU_awareHTN),]
  
  excl$aware <<- nrow(data)
  
  # Only interested in treated hypertensives
  data <- data[data$TEU_treatedHTN == TRUE & !is.na(data$TEU_treatedHTN),]
  
  excl$treated <<- nrow(data)
  
  # Exclude MACE at baseline
  data <- data[!data$TEU_MACE_prev=="Yes",]
  
  excl$MACE <<- nrow(data)
  
  # Restrict to genetically white British
  data <- data[!is.na(data$GeP_ethnic),]
  
  excl$whiteBrit <<- nrow(data)
  
  return(data) 
}

excl <- list(initial=0)

data <-
  evalWithMemoization(
    derive_variables(
      database = "K:/TEU/UKB33952_Data/Data_Downloads/V3_database_duckdb0.2.1/ukb_v3.db",
      field_definitions = TEU_SPECS$HTN_control_MACE,
      exclusions = exclusions
    ),
    key = c(TEU_SPECS$HTN_control_MACE, exclusions)
  )

pretty_func <- pretty_switch(field_definitions=TEU_SPECS$HTN_control_MACE, return_type = "function")
pretty_names <- pretty_switch(field_definitions=TEU_SPECS$HTN_control_MACE, return_type = "list")

export_svg(DiagrammeR::grViz(here::here(file.path(config$outputs$flowcharts, "HTN_PRS_MACE.gv")))
           ) %>% charToRaw %>% rsvg %>% 
  png::writePNG(here::here(file.path(config$outputs$figures, "HTN_PRS_MACE.png")))


```

```{r}
# Temporary fix:

levels(data$TEU_TownsendDepInd_Quint) <- c(levels(data$TEU_TownsendDepInd_Quint), "Unavailable")
data$TEU_TownsendDepInd_Quint[is.na(data$TEU_TownsendDepInd_Quint)] <- "Unavailable"

```


# Hypertension PRS cross-sectional analysis

## Table `r table`. Characteristics of participants included in the analysis, by Systolic BP PRS quintile (treated hypertensives, n = `r nrow(data)`).
`r table <- table + 1`

```{r descriptive-table, include=TRUE, echo=FALSE}
varlist=c("TEU_BaC_AgeAtRec", "TEU_BaC_AgeCat", "BaC_Sex", "TEU_BlP_SBP_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country"
          )

tab_all <- descriptivetable(df=data, 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names)

tab_1 <- descriptivetable(df=data[data$TEU_SBP_PRS_quintiles=="Q1: lowest",], 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names)

tab_3 <- descriptivetable(df=data[data$TEU_SBP_PRS_quintiles=="Q3",], 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names)

tab_5 <- descriptivetable(df=data[data$TEU_SBP_PRS_quintiles=="Q5: highest",], 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names)

tab <- inner_join(tab_all, tab_1%>%select(-Variable), by="IDcol")
tab <- inner_join(tab, tab_3%>%select(-Variable), by="IDcol")
tab <- inner_join(tab, tab_5%>%select(-Variable), by="IDcol")

rownames(tab) <- NULL
tab <- rbind(c("", "Variable", "n", "%", "n", "%", "n", "%", "n", "%"),
              as.matrix(tab))
colnames(tab) <- c("", "", "Overall", "", "Q1: lowest", "", "Q3", "", "Q5: highest", "")

# Output the resulting table
kable(tab[,-1])

```

## Table `r table`: Multivariable logistic regression identifying factors associated with hypertension control, among middle-aged UK adults on anti-hypertensive treatment (n=`r nrow(data)`)
`r table <- table + 1`

```{r logistic-regression, include=TRUE, echo=FALSE}

varlist=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_SBP_PRS_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

desctable2 <- function(df){
  n <- nrow(df)
  control <- prop.test(x=nrow(df[df$TEU_controlledHTN==TRUE & !is.na(df$TEU_controlledHTN),]),
                           n=nrow(df), alternative="two.sided", conf.level=0.95)
  pct_control <- pretty_dp(control$estimate, dp=1, pct=TRUE)
  
  table <- cbind(n, pct_control)
  return(table)
}

pct_ctrl <- c()
for(var in varlist){
  header <- c(var, paste0("**", pretty_func(var), "**"), NA, NA)
  pct_ctrl <- rbind(pct_ctrl, header)
  for(level in levels(data[[var]])){
    newrow <- desctable2(data[data[var]==level,])
    IDcol <- paste0(var, level)
    pct_ctrl <- rbind(pct_ctrl, cbind(IDcol, level, newrow))
  }
}
pct_ctrl <- data.frame(pct_ctrl)

adj <- regressiontable(df=data, outcome="TEU_controlledHTN", varlist=varlist, 
                       regresstype="adjusted", adjvarlist=c("TEU_BaC_AgeCat", "BaC_Sex"))
multi <- regressiontable(df=data, outcome="TEU_controlledHTN", varlist=varlist, 
                         regresstype="multivariable")

tab3 <- merge(adj, multi, by=c("IDcol", "Coefficient", "Levels"), sort=FALSE) %>% select(-Coefficient, -Levels)
# Add n column

tab3 <- left_join(pct_ctrl, tab3, by=c("IDcol")) %>% select(-IDcol)

# Hide "unanswered"/"unknown"/"do not know" categories
# tab3 <- tab3[-which(tab3$level %in% c("Unknown", "Unanswered", "Do not know", "Prefer not to answer")),]
rownames(tab3) <- NULL

tab3 <- rbind(c("Coefficient", "n^1^", "% Controlled^2^", "OR", "95% CI", "p", "OR", "95% CI", "p"),
              as.matrix(tab3))
colnames(tab3) <- c("","","", 
                    "Age- and sex-adjusted", "", "", 
                    "Multivariable- adjusted", "", "")

# Output the resulting table
kable(tab3)

```

# Hypertension PRS prospective analysis

Including SBP PRS (quintiles), showing for each outcome:

* Model A: without measured SBP at baseline
* Model B: with measured SBP at baseline

## Cox regression of time until MACE event (`r nrow(data[data$TEU_MACE_status==1,])` events)

MI event defined as diagnosis of MACE event or MACE prevention procedure in HES data, or MACE recorded as primary cause of death.

```{r Cox model mace, echo=FALSE}

varlist1=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_SBP_PRS_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

varlist2=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_SBP_PRS_quintiles", "TEU_BlP_SBP_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

formula1 <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_status) ~ ", paste(varlist1, collapse="+")))

rmodel1<-coxph(formula1,data=data)

routput1 <- publish(rmodel1, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable %>%
  select(-signif) %>%
  mutate(ID = ifelse(Variable=="", NA, Variable)) %>%
  tidyr::fill(ID)

formula2 <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_status) ~ ", paste(varlist2, collapse="+")))

rmodel2<-coxph(formula2,data=data)

routput2 <- publish(rmodel2, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable %>%
  select(-signif) %>%
  mutate(ID = ifelse(Variable=="", NA, Variable)) %>%
  tidyr::fill(ID)


tab4 <- left_join(routput2, routput1, by=c("ID", "Variable", "Units"), suffix=c("_b", "_a")) %>%
  select(Variable, Units, HazardRatio_a, CI.95_a, `p-value_a`, HazardRatio_b, CI.95_b, `p-value_b`) %>%
  replace(., is.na(.), "")

# Change pretty variable names
tab4$Variable <-  sapply(tab4$Variable,function(i) if(i!=""){pretty_func(i)}else{i})

rownames(tab4) <- NULL
tab4 <- rbind(c("Coefficient", "Level", "OR", "95% CI", "p", "OR", "95% CI", "p"),
              as.matrix(tab4))
colnames(tab4) <- c("","", 
                    "Model A", "", "", 
                    "Model B", "", "")

kable(tab4)


```

## Cox regression of time until MI event (`r nrow(data[data$TEU_MACE_MI==1,])` events)

MI event defined as diagnosis of myocardial infarction or MI prevention procedure in HES data, or MI recorded as primary cause of death (censoring at first MACE event applied)

```{r Cox model MI, echo=FALSE}

formula1 <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_MI) ~ ", paste(varlist1, collapse="+")))

rmodel1<-coxph(formula1,data=data)

routput1 <- publish(rmodel1, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable %>%
  select(-signif) %>%
  mutate(ID = ifelse(Variable=="", NA, Variable)) %>%
  tidyr::fill(ID)

formula2 <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_MI) ~ ", paste(varlist2, collapse="+")))

rmodel2<-coxph(formula2,data=data)

routput2 <- publish(rmodel2, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable %>%
  select(-signif) %>%
  mutate(ID = ifelse(Variable=="", NA, Variable)) %>%
  tidyr::fill(ID)


tab4 <- left_join(routput2, routput1, by=c("ID", "Variable", "Units"), suffix=c("_b", "_a")) %>%
  select(Variable, Units, HazardRatio_a, CI.95_a, `p-value_a`, HazardRatio_b, CI.95_b, `p-value_b`) %>%
  replace(., is.na(.), "")

# Change pretty variable names
tab4$Variable <-  sapply(tab4$Variable,function(i) if(i!=""){pretty_func(i)}else{i})

rownames(tab4) <- NULL
tab4 <- rbind(c("Coefficient", "Level", "OR", "95% CI", "p", "OR", "95% CI", "p"),
              as.matrix(tab4))
colnames(tab4) <- c("","", 
                    "Model A", "", "", 
                    "Model B", "", "")

kable(tab4)

```

## Cox regression of the time until stroke event (`r nrow(data[data$TEU_MACE_Stroke==1,])` events)

Stroke event defined as diagnosis of stroke (any kind) or stroke prevention procedure in HES data, or stroke recorded as primary cause of death (censoring at first MACE event applied).

```{r Cox model stroke, echo=FALSE}

formula1 <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_Stroke) ~ ", paste(varlist1, collapse="+")))

rmodel1<-coxph(formula1,data=data)

routput1 <- publish(rmodel1, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable %>%
  select(-signif) %>%
  mutate(ID = ifelse(Variable=="", NA, Variable)) %>%
  tidyr::fill(ID)

formula2 <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_Stroke) ~ ", paste(varlist2, collapse="+")))

rmodel2<-coxph(formula2,data=data)

routput2 <- publish(rmodel2, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable %>%
  select(-signif) %>%
  mutate(ID = ifelse(Variable=="", NA, Variable)) %>%
  tidyr::fill(ID)


tab4 <- left_join(routput2, routput1, by=c("ID", "Variable", "Units"), suffix=c("_b", "_a")) %>%
  select(Variable, Units, HazardRatio_a, CI.95_a, `p-value_a`, HazardRatio_b, CI.95_b, `p-value_b`) %>%
  replace(., is.na(.), "")

# Change pretty variable names
tab4$Variable <-  sapply(tab4$Variable,function(i) if(i!=""){pretty_func(i)}else{i})

rownames(tab4) <- NULL
tab4 <- rbind(c("Coefficient", "Level", "OR", "95% CI", "p", "OR", "95% CI", "p"),
              as.matrix(tab4))
colnames(tab4) <- c("","", 
                    "Model A", "", "", 
                    "Model B", "", "")

kable(tab4)

```

