---
title: "HTN PRS MACE prospective"
author:  "Jennifer Collister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  word_document:
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)
library(Publish)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)
library(png)
library(dplyr)
library(ggplot2)
library(survival)
library(pROC)

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(config$functions$JC))

#-------------------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r data-derivation, include=FALSE, cache=TRUE}
 
source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data){
  
  excl$initial <<- nrow(data)
  
  # Exclude those outside the 40-70 age range
  data <- data[data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  
  excl$agerange <<- nrow(data)

  # Exclude individuals with missing BP data
  # or missing answers to BP questions on touchscreen questionnaire
  data <- data[!is.na(data$TEU_BlP_SBP.avg),]
  data <- data[!is.na(data$TEU_BlP_DBP.avg),]
  data <- data[data$TEU_HMH_prevHTN != "Unanswered",]
  data <- data[data$TEU_HMH_Meds_BP != "Unanswered",]
  # Exclude participants who only had BP measured once
  data <- data[data$TEU_BlP_nSBP == 2 & data$TEU_BlP_nDBP == 2,]
  
  excl$BPmiss <<- nrow(data)

  # Exclude individuals with implausible BP data
  data <- data[data$TEU_BlP_SBP.avg >= 70 & data$TEU_BlP_SBP.avg <= 270,]
  data <- data[data$TEU_BlP_DBP.avg >= 50 & data$TEU_BlP_DBP.avg <= 150,]
  
  excl$BPimp <<- nrow(data)
  
  # Only keep those with genetic data
  data <- data[!is.na(data$TEU_BP_PRS),]
  
  excl$PRSmiss <<- nrow(data)

  # Exclude pregnant women ("Yes" or "Unsure")
  data <- data[is.na(data$VeI_PregnantNow) | data$VeI_PregnantNow == "No",]
  
  excl$pregnant <<- nrow(data)

  # Exclude individuals who have serious health conditions
  data <- data[is.na(data$TEU_VeI_seriouscomb),]

  # And individuals with cancer (except for skin cancer?)
  data <- data[is.na(data$TEU_VeI_cancer),]
  
  excl$seriouscomorb <<- nrow(data)
  
  # Any durations less than 0 are clearly errors (only 2)
  # Any diagnoses before the age of about 20 are probably due to other causes
  data$TEU_VeI_HTN_dur[data$TEU_VeI_HTN_dur < 0] <- NA
  # data[data$TEU_VeI_HTN_dur > data$TEU_BaC_AgeAtRec-20,]
  
  # Exclude those who were not hypertensive
  data <- data[data$TEU_evidenceHTN == TRUE & !is.na(data$TEU_evidenceHTN),]
  
  excl$evidenceHTN <<- nrow(data)
  
  # Exclude those who were not aware
  data <- data[data$TEU_awareHTN == TRUE & !is.na(data$TEU_awareHTN),]
  
  excl$aware <<- nrow(data)
  
  # Only interested in treated hypertensives
  data <- data[data$TEU_treatedHTN == TRUE & !is.na(data$TEU_treatedHTN),]
  
  excl$treated <<- nrow(data)
  
  # Exclude MACE at baseline
  data <- data[!data$TEU_MACE_prev=="Yes",]
  
  excl$MACE <<- nrow(data)
  
  # Restrict to genetically white British
  data <- data[!is.na(data$GeP_ethnic),]
  
  excl$whiteBrit <<- nrow(data)
  
  return(data)
}

excl <- list(initial=0)

data <-
  evalWithMemoization(
    derive_variables(
      database = "K:/TEU/UKB33952_Data/Data_Downloads/V3_database_duckdb0.2.1/ukb_v3.db",
      field_definitions = TEU_SPECS$HTN_control_MACE,
      exclusions = exclusions
    ),
    key = c(TEU_SPECS$HTN_control_MACE, exclusions)
  )

pretty_func <- pretty_switch(field_definitions=TEU_SPECS$HTN_control_MACE, return_type = "function")
pretty_names <- pretty_switch(field_definitions=TEU_SPECS$HTN_control_MACE, return_type = "list")

export_svg(DiagrammeR::grViz(here::here(file.path(config$outputs$flowcharts, "HTN_PRS_MACE.gv")))
           ) %>% charToRaw %>% rsvg %>% 
  png::writePNG(here::here(file.path(config$outputs$figures, "HTN_PRS_MACE.png")))


```

# Analysis Population

## Figure `r figure`. Flowchart illustrating selection of analytic dataset (n = `r nrow(data)`).
`r figure <- figure + 1`
![Missing BP data was fewer than 2 BP measurements, or missing answers to the prior diagnosis and treatment questions. Hypertension was defined as self-report of hypertensive medication use, or self-report of a prior diagnosis of hypertension, or mean BP >= 140 /90 mmHg at baseline assessment. Awareness was defined as self-report of a prior diagnosis of hypertension among those who were hypertensive. Treatment was defined as self-report of hypertensive medication among those who were aware. Control was defined as mean BP < 140/90 mmHg at baseline assessment, among those who were treated.](Figures/HTN_PRS_MACE.png){height=700px}

# Hypertension PRS cross-sectional analysis

## Table `r table`. Characteristics of participants included in the analysis, by BP PRS quintile (treated hypertensives, n = `r nrow(data)`).
`r table <- table + 1`
```{r descriptive-table, include=TRUE, echo=FALSE}
varlist=c("TEU_BaC_AgeAtRec", "TEU_BaC_AgeCat", "BaC_Sex",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country"
          )

footnote_list <- c("TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country", "TEU_BSM_BMIcat", "Prosp_comorb_numcat")
footnote_no <- 1

tab_all <- descriptivetable(df=data, 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names,
                         footnote_list=footnote_list)

tab_1 <- descriptivetable(df=data[data$TEU_BP_PRS_quintiles=="Q1: lowest",], 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names,
                         footnote_list=footnote_list)

tab_3 <- descriptivetable(df=data[data$TEU_BP_PRS_quintiles=="Q3",], 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names,
                         footnote_list=footnote_list)

tab_5 <- descriptivetable(df=data[data$TEU_BP_PRS_quintiles=="Q5: highest",], 
                         varlist=varlist,
                         contavg='median',
                         pretty_names=pretty_names,
                         footnote_list=footnote_list)

tab <- inner_join(tab_all, tab_1%>%select(-Variable), by="IDcol")
tab <- inner_join(tab, tab_3%>%select(-Variable), by="IDcol")
tab <- inner_join(tab, tab_5%>%select(-Variable), by="IDcol")

rownames(tab) <- NULL
tab <- rbind(c("", "Variable", "n", "%", "n", "%", "n", "%", "n", "%"),
              as.matrix(tab))
colnames(tab) <- c("", "", "Overall", "", "Q1: lowest", "", "Q3", "", "Q5: highest", "")

# Output the resulting table
add_footnote(kable(tab[,-1]), label=c("Occupation categories have been condensed from those recorded in UKB. Professional and Administrative: Managers and Senior Officials, Professional Occupations, Associate Professional and Technical Occupations, Administrative and Secretarial Occupations. Skilled trades: Skilled Trades Occupations. Services: Personal Service Occupations, Sales and Customer Service Occupations. Manual and Industrial: Process, Plant and Machine Operatives, Elementary Occupations. Other employment: free text entry that was not coded by UKB.", 
                                  "Self-reported highest education achieved was mapped to the International Standard Classification of Education (ISCED) categories",
                                  "The assessment centre the participant attended was used as a proxy for country of residence",
                                  "BMI has been categorised as: Underweight < 18.5 kg/m^2^; Normal 18.5 - 24.9 kg/m^2^; Overweight 25.0 - 29.9 kg/m^2^; Obese >= 30.0 kg/m^2^.", 
                                  "The number of comorbidities considered to be of interest - diabetes, arrhythmia (afib/flutter), asthma or COPD, migraines, epilepsy, anxiety, depression, osteoarthritis, other joint disorder."), notation="number")

```


## Table `r table`: Multivariable logistic regression identifying factors associated with hypertension control, among middle-aged UK adults on anti-hypertensive treatment (n=`r nrow(data)`)
`r table <- table + 1`
<div custom-style="Todo">Reminder: Manually re-order BMI categories</div>

```{r logistic-regression, include=TRUE, echo=FALSE}
library(forestplot)
varlist=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_BP_PRS_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

# formula <- paste0("TEU_controlledHTN ~ ", paste(varlist, collapse=" + "))
formula <- as.formula(paste0("TEU_controlledHTN ~ ", paste(varlist, collapse="+")))

agesex <- glm("TEU_controlledHTN ~ TEU_BaC_AgeCat + BaC_Sex + TEU_BP_PRS_quintiles", data=data, family="binomial")
multi <- glm(formula, data=data, family="binomial")


summ <- summary(agesex)
conf <- summ$coeff
conf[,3] <- conf[,1] - (1.96*conf[,2])
conf[,4] <- conf[,1] + (1.96*conf[,2])
unctrl <- data.frame(hr = c(0, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),1]),
lci = c(0, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),3]),
uci = c(0, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),4])
)

summ <- summary(multi)
conf <- summ$coeff
conf[,3] <- conf[,1] - (1.96*conf[,2])
conf[,4] <- conf[,1] + (1.96*conf[,2])
ctrl <- data.frame(hr = c(0, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),1]),
lci = c(0, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),3]),
uci = c(0, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),4])
)

# tabletext <- 
tabletext <- c("Age- and sex-adjusted", "Multivariate")

forestplot(tabletext,
           legend = c("Q1 (lowest)", "Q2", "Q3", "Q4", "Q5 (highest)"),
           legend_args = fpLegend(title="Quintiles of BP PRS",
                                  gp = gpar(col = "#CCCCCC", fill = "#F9F9F9")),
           mean = t(cbind(unctrl$hr, ctrl$hr)),
           lower = t(cbind(unctrl$lci, ctrl$lci)),
           upper = t(cbind(unctrl$uci, ctrl$uci)),
           col = fpColors(box = c("green", "blue", "purple", "red", "orange")),
           xlab = "OR for hypertension control",
           vertices=TRUE
           )

```

```{r logistic-regression, include=TRUE, echo=FALSE}
library(forestplot)
varlist=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_BP_PRS_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

# formula <- paste0("TEU_controlledHTN ~ ", paste(varlist, collapse=" + "))
formula <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_status) ~ ", paste(varlist, collapse="+")))

uncontrol <- coxph(formula, data=data[data$TEU_controlledHTN==FALSE,])
control <- coxph(formula, data=data[data$TEU_controlledHTN==TRUE,])
all <- coxph(formula, data=data)
# all <- glm(formula, data=data, family="binomial")

# tabletext <- c("1st quintile", "2nd", "3rd quintile", "4th", "5th quintile")
tabletext <- c("Uncontrolled", "Controlled", "Overall")

summ <- summary(uncontrol)
conf <- summ$conf.int
unctrl <- data.frame(hr = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),1]),
lci = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),3]),
uci = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),4])
)

summ <- summary(control)
conf <- summ$conf.int
ctrl <- data.frame(hr = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),1]),
lci = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),3]),
uci = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),4])
)

summ <- summary(all)
conf <- summ$conf.int
all <- data.frame(hr = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),1]),
lci = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),3]),
uci = c(1, conf[rownames(conf) %in% c("TEU_BP_PRS_quintilesQ2", "TEU_BP_PRS_quintilesQ3", 
                                      "TEU_BP_PRS_quintilesQ4", "TEU_BP_PRS_quintilesQ5: highest"),4])
)

forestplot(tabletext,
       mean = t(cbind(unctrl$hr, ctrl$hr,all$hr)),
       lower = t(cbind(unctrl$lci, ctrl$lci, all$lci)),
       upper = t(cbind(unctrl$uci, ctrl$uci, all$uci)),
       )

# ggforest(model)



```


# Hypertension PRS prospective analysis

## Total MACE outcomes by subtype

```{r MACE subtypes,echo=FALSE}

# Freq of MACE subtypes
MACE_subtypes <- data.frame(propped(table(data$TEU_MACE_fucomp)))
colnames(MACE_subtypes)[1] <- 'MACE subtypes'

kable(MACE_subtypes[-nrow(MACE_subtypes),],caption=paste0('Table.1 Frequency of MACE subtypes (N=',MACE_subtypes[nrow(MACE_subtypes),2],')'))

  
```

## Cox regression of time until MACE event (`r nrow(data[!is.na(data$TEU_MACE_fucomp),])` events)


```{r Cox model mace, echo=FALSE}

varlist=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_BP_PRS_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

# age- and sex-adjusted
agesex <- coxph(Surv(TEU_MACE_time_yrs,TEU_MACE_status) ~ TEU_BaC_AgeCat + BaC_Sex, data=data)
adj <- publish(agesex, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable
for(var in varlist){
  if(! var %in% c("TEU_BaC_AgeCat", "BaC_Sex")){
  formula <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_status) ~ TEU_BaC_AgeCat + BaC_Sex + ", var))
  model <- coxph(formula, data=data)
  output <- publish(model, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable
  output <- output[-c(1:5),] %>% select(-any_of("Missing"))
  adj <- rbind(adj, output)
  }
}

# Complete case data
formula <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_status) ~ ", paste(varlist, collapse="+")))

rmodel1<-coxph(formula,data=data)


routput1 <- publish(rmodel1, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable  %>%
  select(-any_of("Missing"))

tab4 <- cbind(adj%>%select(-signif), routput1%>%select(-Variable, -Units, -signif))

# Change pretty variable names
tab4$Variable <-  sapply(tab4$Variable,function(i) if(i!=""){pretty_func(i)}else{i})

rownames(tab4) <- NULL
tab4 <- rbind(c("Coefficient", "Level", "OR", "95% CI", "p", "OR", "95% CI", "p"),
              as.matrix(tab4))
colnames(tab4) <- c("","", 
                    "Age- and sex-adjusted", "", "", 
                    "Multivariable- adjusted", "", "")

kable(tab4)


```

## Cox regression of the time until haemorrhagic stroke event (`r nrow(data[data$TEU_MACE_Stroke==1,])` events)

Haemorrhagic stroke event defined as diagnosis of haemorrhagic stroke in HES data or haemorrhagic stroke recorded as primary cause of death (censoring at first MACE event applied).

```{r Cox model stroke, echo=FALSE}

varlist=c("TEU_BaC_AgeCat", "BaC_Sex", "TEU_BP_PRS_quintiles",
          "TEU_BSM_BMIcat", "TEU_Smo_Status", "TEU_Alc_WeeklyCat", "TEU_Pha_METsover1200",
          "TEU_FaH_CVD", "Prosp_comorb_numcat", "TEU_VeI_numHTNmedscat", "TEU_LDL_Quintiles",
          "TEU_TownsendDepInd_Quint", "TEU_HouseholdIncome", "TEU_Emp_category", "TEU_Edu_ISCED", "TEU_Rec_Country",
          "GeP_PC_1", "GeP_PC_2", "GeP_PC_3", "GeP_PC_4"
          )

# age- and sex-adjusted
agesex <- coxph(Surv(TEU_MACE_time_yrs,TEU_MACE_Stroke) ~ TEU_BaC_AgeCat + BaC_Sex, data=data)
adj <- publish(agesex, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable
for(var in varlist){
  if(! var %in% c("TEU_BaC_AgeCat", "BaC_Sex")){
  formula <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_Stroke) ~ TEU_BaC_AgeCat + BaC_Sex + ", var))
  model <- coxph(formula, data=data)
  output <- publish(model, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable
  output <- output[-c(1:5),] %>% select(-any_of("Missing"))
  adj <- rbind(adj, output)
  }
}

# Complete case data
formula <- as.formula(paste0("Surv(TEU_MACE_time_yrs,TEU_MACE_Stroke) ~ ", paste(varlist, collapse="+")))

rmodel1<-coxph(formula,data=data)


routput1 <- publish(rmodel1, ci.digits=2, ci.handler="sprintf", pvalue.eps=0.01,
              pvalue.digits=2,pvalue.stars=TRUE, ci.trim=FALSE,print=FALSE)$regressionTable  %>%
  select(-any_of("Missing"))

tab4 <- cbind(adj%>%select(-signif), routput1%>%select(-Variable, -Units, -signif))

# Change pretty variable names
tab4$Variable <-  sapply(tab4$Variable,function(i) if(i!=""){pretty_func(i)}else{i})

rownames(tab4) <- NULL
tab4 <- rbind(c("Coefficient", "Level", "OR", "95% CI", "p", "OR", "95% CI", "p"),
              as.matrix(tab4))
colnames(tab4) <- c("","", 
                    "Age- and sex-adjusted", "", "", 
                    "Multivariable- adjusted", "", "")

kable(tab4)


```
