---
title: "BP PRS Analysis"
author: "Jennifer Collister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  word_document:
      reference_docx: "K:/TEU/TEU Guides/TEU_DocStyle_2020.dotx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here())

# Load the project config file for filepaths etc
config = yaml.load_file(here("./config.yml"))

# Load data
data <- readRDS(file=paste0(config$analysisdata$htn, "HTN_excl.rds"))
prs <- readRDS(file=paste0(config$analysisdata$prs, "Evangelou2018_PRS.rds"))


```

1. Confirm the overall PRS on the whole data set match what Jemma’s group get – at least to the first decimal place.  If not, explore why.

Waiting to hear back from Jemma's team

2. Table 1 analysis.  Construct a Table 1a.  Horizontal Quintiles 1, 3, 5 of SBP PRS.  Table 1b: Same for DBP PRS.  Vertical risk factors: As for Table 1 in HTNTables_JC20200624.doc.    If there are non-trivial differences by age and/or sex will need to standardize the other risk factors (unlikely to be needed).

```{r}


```

3. Table 2 analysis.  Add the SBP PRS and DBP PRS individually, and together (i.e. 3 multivariate analyses).

```{r}


```

4. Compute observed SBP and DBP by quintile of each PRS (watchout for influence of outliers).  A box and whisker plot would be good.

```{r}


```